---
title: "RFID Room 3"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

# Data Preprocessing

## Function to find the most frequent zone (average wont quite work)

```{r find most frequent zone}
MaxTable <- function(x){
     dd <- unique(x)
     dd[which.max(tabulate(match(x,dd)))]
}

```

## My to minute function

```{r my to minute}
to_minute <- function(xts_object){
  ep_min <- endpoints(na.omit(xts_object), on="minutes",k=60)
  ave_min <- period.apply(na.omit(xts_object), INDEX=ep_min, FUN=MaxTable)
  #this removes the time to allow cbinding
  ave_min <- to.minutes(ave_min,OHLC=F,indexAt="startof")
  return(ave_min)
}
```

## My to daily function

```{r my to daily ignore}
# to_daily <- function(xts_object){
#   ep_day <- endpoints(na.omit(xts_object), on="days",k=1)
#   ave_day <- period.apply(na.omit(xts_object), INDEX=ep_day, FUN=MaxTable)
#   #this removes the time to allow cbinding
#   ave_day <- to.daily(ave_day,OHLC=F,indexAt="startof")
#   return(ave_day)
# }
```

```{r my to daily}
 to_daily <- function(xts_object){
   ep_day <- endpoints(na.omit(xts_object), on="minutes",k=1440)
   ave_day <- period.apply(na.omit(xts_object), INDEX=ep_day, FUN=MaxTable)
   #this removes the time to allow cbinding
   ave_day <- to.daily(ave_day,OHLC=F,indexAt="startof")
   return(ave_day)
 }
```

## My to weekly function

```{r my to weekly}
to_weekly <- function(xts_object){
  ep_week <- endpoints(na.omit(xts_object), on="minutes",k=10080)
  ave_week <- period.apply(na.omit(xts_object), INDEX=ep_week, FUN=MaxTable)
  #this removes the time to allow cbinding
  ave_week <- to.weekly(ave_week,OHLC=F,indexAt="startof")
  return(ave_week)
}
```

## My to monthly function

```{r my to monthly}
to_monthly <- function(xts_object){
  ep_month <- endpoints(na.omit(xts_object), on="minutes",k=43800)
  ave_month <- period.apply(na.omit(xts_object), INDEX=ep_month, FUN=MaxTable)
  #this removes the time to allow cbinding
  ave_month <- to.monthly(ave_month,OHLC=F,indexAt="startof")
  return(ave_month)
}
```


## Function to Calculate Transitions

```{r function to calc transitions}
#I don't know when we use this data yet, will be useful when making transition tables for the whole dataset 
sep_bird_id_xts <- function(samp_name, samp_id,cage_obj){
  
  raw_sample_tab <- subset(cage_obj, tagname==samp_id)
  
  xts_object <- xts(raw_sample_tab ,order.by = raw_sample_tab$DateTime)
  
  result <- list("name" = samp_name, "ID"=samp_id, "xts_obj"=xts_object)
  return(result)
}

sep_bird_id_period <- function(samp_name, samp_id, cage_obj, cutoff){
  
  raw_sample_tab <- subset(cage_obj, tagname==samp_id)
  
  xts_object <- xts(raw_sample_tab ,order.by=raw_sample_tab$DateTime)
  
  xts_object <- xts_object[cutoff]
  
  daily <- to_daily(xts_object$subzone)#, OHLC=F)
  weekly <- to_weekly(xts_object$subzone)#, OHLC=F)
  monthly <- to_monthly(xts_object$subzone)#, OHLC=F)
  index(monthly) <-as.POSIXct(index(monthly))
  
  result <- list("name" = samp_name, "ID"=samp_id, "xts_obj"=xts_object, "daily_obj"=daily,
                 "weekly_obj"=weekly, "monthly_obj"=monthly)
  return(result)
}

#input a sample name
#returns a vector labeled with sample name, and bottom, middle, top and total transitions

calc_trans<- function(samp_name,samp_obj, id){
  bottom_trans <-0
  mid_trans<-0
  top_trans<-0
  trans<-0
  count <- 0
  
  for (i in 2:length(samp_obj$subzone)) {
  previous_state <- as.character(samp_obj$subzone[i-1])
  current_state <- as.character(samp_obj$subzone[i])
  
  #print(paste(previous_state,":",current_state))
  
  
  if (previous_state == current_state) {
    count <- count+1
  }
  else{
    
    if((current_state == "bottom") || (current_state == "Bottom")){
      bottom_trans <- bottom_trans+1
      trans<- trans+1
    }
    if (current_state=="middle" || current_state == "Middle") {
      mid_trans <- mid_trans+1
      trans <- trans+1
    }
    if(current_state =="top" || current_state=="Top"){
      top_trans <- top_trans+1
      trans <- trans+1
    }
  }
  }
  
  result <- c(samp_name,id,bottom_trans,mid_trans,top_trans,trans)
  return(result)
  
}

calc_trans_period<- function(samp_name, daily_indexed, raw_table,freq){
  result <- data.frame()
  
  for (i in 2:(length(daily_indexed))){
    curr_day <- raw_table[paste0(index(daily_indexed)[i-1],"/",index(daily_indexed)[i])]
    curr_day_trans <- calc_trans(paste0(samp_name,".",freq,".",i),curr_day, samp_name)
    result <- rbind(result,curr_day_trans)
    }
  colnames(result) <- c("sample","ID","bottom","mid","top","total")
  return(result)
}
```

```{r my calc trans duration function}

calc_trans_duration<- function(samp_name,samp_obj, id){
  bottom_trans <-0
  mid_trans<-0
  top_trans<-0
  trans<-0
  count <- 0
  bottom_time <- 0 
  mid_time <- 0
  top_time <- 0
  
  for (i in 2:length(samp_obj$subzone)) {
  previous_state <- as.character(samp_obj$subzone[i-1])
  current_state <- as.character(samp_obj$subzone[i])
  previous_time <- index(samp_obj[i-1])
  current_time <- index(samp_obj[i])
  
  #print(paste(previous_state,":",current_state))
  
  
  if (previous_state == current_state) {
    count <- count+1
  }
  else{
    
    if((previous_state == "bottom") || (previous_state == "Bottom")){
      bottom_trans <- bottom_trans+1
      trans<- trans+1
      bottom_time <- bottom_time + difftime(current_time, previous_time,units="secs")
    }
    if (previous_state=="middle" || previous_state == "Middle") {
      mid_trans <- mid_trans+1
      trans <- trans+1
      mid_time <- mid_time + difftime(current_time, previous_time, units="secs")
    }
    if(previous_state =="top" || previous_state=="Top"){
      top_trans <- top_trans+1
      trans <- trans+1
      top_time <- top_time + difftime(current_time,previous_time,units="secs")
    }
  }
  }
  
  result <- c(samp_name,id,bottom_trans,mid_trans,top_trans,trans,bottom_time,mid_time,top_time)
  #colnames(result) <- c("sample_name","ID","bottom_trans","middle_trans","top_trans","total_trans","bottom_sec","mid_sec","top_sec")
  return(result)
  
}

calc_trans_period_duration<- function(samp_name, daily_indexed, raw_table,freq){
  result <- data.frame()
  
  for (i in 2:(length(daily_indexed))){
    curr_day <- raw_table[paste0(index(daily_indexed)[i-1],"/",index(daily_indexed)[i])]
    curr_day_trans <- calc_trans_duration(paste0(samp_name,".",freq,".",i),curr_day, samp_name)
    result <- rbind(result,curr_day_trans)
    }
  colnames(result) <- c("sample_name","ID","bottom_trans","middle_trans","top_trans","total_trans","bottom_sec","mid_sec","top_sec")
  return(result)
}
```

## Import 9013

```{r import 9013}
library(xts)
#65495
nine_zero_one_three <- read.csv("../data/9013.csv")

bird_ids_9013 <- unique(nine_zero_one_three$tagname)
bird_ids_9013 <- na.trim(sort(bird_ids_9013))

nine_zero_one_three["DateTime"] <- as.POSIXct(nine_zero_one_three$access, origin="1970-01-01", tz="GMT")

print("what makes up subzone col")
unique(nine_zero_one_three$subzone)

nine_zero_one_three$subzone[nine_zero_one_three$subzone == "Bottom"] <- "bottom"
nine_zero_one_three$subzone[nine_zero_one_three$subzone == "Middle"] <- "middle"
nine_zero_one_three$subzone[nine_zero_one_three$subzone == "Top"] <- "top"

#65495
print("what makes up subzone col")
unique(nine_zero_one_three$subzone)

print("how many NA's are in Datetime and Subzone")
sum(is.na(nine_zero_one_three$DateTime))
sum(is.na(nine_zero_one_three$subzone))


#generate list of XTS objects 
reslist_9013 <- list()
for(i in 1:length(bird_ids_9013)){
  
  res <- sep_bird_id_xts(samp_name="9013",cage_obj =nine_zero_one_three, samp_id = bird_ids_9013[i])
  reslist_9013[[i]] <- res
}



remove <- c("6998")
bird_ids_9013_new <- bird_ids_9013 [! bird_ids_9013  %in% remove ]
#generate list of the first 10 truncated XTS objects
reslist2_9013 <- list()
for(i in 1:length(bird_ids_9013_new)){
  res <- sep_bird_id_period(samp_name="9013",cage_obj =nine_zero_one_three, samp_id = bird_ids_9013_new[i],cutoff = "2021-03-09 T18:00:00/2021-05-07 T17:00:00")
  reslist2_9013[[i]] <- res
}



#find wholly non-na daily timeset, to be able to feed into the transcalc
big_table_9013 <- reslist2_9013[[1]]$daily_obj
for(i in 2:length(reslist2_9013)){
  current <- reslist2_9013[[i]]$daily_obj
  big_table_9013 <- cbind(big_table_9013,current)
}
print("how many NA's are in big_table_9013")
print(sum(is.na(big_table_9013)))
```

# Intra-Bird Analysis

```{r }
whole_data_sanity <- data.frame()
whole_data_sanity <- reslist2_9013[[1]]$daily_obj
for(item in reslist2_9013){
  whole_data_sanity <- cbind(whole_data_sanity, item$daily_obj)
}
print("how many NAs appear in the full dataset")
sum(is.na(whole_data_sanity))

trans_reslist_room_3 <- list()

for(i in 1:length(reslist2_9013)){
  
  daily_trans_table <- calc_trans_period(reslist2_9013[[i]]$ID,reslist2_9013[[i]]$daily_obj,reslist2_9013[[i]]$xts_obj,"d")
  weekly_trans_table <- calc_trans_period(reslist2_9013[[i]]$ID,reslist2_9013[[i]]$weekly_obj,reslist2_9013[[i]]$xts_obj,"w")
  monthly_trans_table <- calc_trans_period(reslist2_9013[[i]]$ID,reslist2_9013[[i]]$monthly_obj,reslist2_9013[[i]]$xts_obj,"m")
    
  result <- list("ID"=reslist2_9013[[i]]$ID,"daily"=daily_trans_table, "weekly"=weekly_trans_table,"monthly"=monthly_trans_table, "raw"=reslist2_9013[[i]]$xts_obj)
  trans_reslist_room_3[[i]] <- result
}

room_3_daily <- data.frame()
room_3_weekly <- data.frame()
room_3_monthly <- data.frame()
#TODO select sample as opposed to ID for intra bird comparison (alpha eq)
for(item in trans_reslist_room_3){
  room_3_daily <- rbind(room_3_daily, item$daily[c("sample","bottom","mid","top","total")])
  room_3_weekly <- rbind(room_3_weekly, item$weekly[c("sample","bottom","mid","top","total")])
  room_3_monthly <- rbind(room_3_monthly, item$monthly[c("sample","bottom","mid","top","total")])
}

# check to make sure only one bird is selected
print("check that the id and RFID tag are unique for each item entry")
#this is following up with the issues of the intra-bird comparisons
for(item in trans_reslist_room_3){
  print(paste(unique(item$raw$Tagnumber)," ",unique(item$raw$tagname)))
  #print(item)
}



write.csv(room_3_daily,"../output/room_3_daily_intra.csv",row.names = F)
write.csv(room_3_weekly,"../output/room_3_weekly_intra.csv",row.names = F)
write.csv(room_3_monthly,"../output/room_3_monthly_intra.csv",row.names = F)

```

# Room 3 hourly dataset

```{r room 3 hourly dataset}

trans_reslist_room3_min <- list()

for(i in 1:length(reslist2_9013)){
  minute_period <- to_minute(reslist2_9013[[i]]$xts_obj$subzone)
  minute_trans_table <- calc_trans_period(reslist2_9013[[i]]$ID, minute_period, reslist2_9013[[i]]$xts_obj, "Hr")  
  result <- list("ID"=reslist2_9013[[i]]$ID,"hour"=minute_trans_table, "raw"=reslist2_9013[[i]]$xts_obj)
  trans_reslist_room3_min[[i]] <- result
}


room3_hour <- data.frame()

for(item in trans_reslist_room3_min){
  room3_hour <- rbind(room3_hour, item$hour[c("sample","bottom","mid","top","total")])
}

# check to make sure only one bird is selected
print("check that the id and RFID tag are unique for each item entry")
#this is following up with the issues of the intra-bird comparisons
for(item in trans_reslist_room3_min){
  print(paste(unique(item$raw$Tagnumber)," ",unique(item$raw$tagname)))
  #print(item)
}
```

## Determine which bird is the most active in the Room 3 Dataset

```{r who is the most active?}
room3_daily <- data.frame()
room3_weekly <- data.frame()
room3_monthly <- data.frame()
#TODO select sample as opposed to ID for intra bird comparison (alpha eq)
for(item in trans_reslist_room_3){
  room3_daily <- rbind(room3_daily, item$daily[c("ID","bottom","mid","top","total")])
  room3_weekly <- rbind(room3_weekly, item$weekly[c("ID","bottom","mid","top","total")])
  room3_monthly <- rbind(room3_monthly, item$monthly[c("ID","bottom","mid","top","total")])
}
tmp <- cbind(as.character(room3_daily$ID), as.numeric(room3_daily$total))
#nrow=32 comes from 32 birds used
tmp2 <- matrix(as.vector(as.numeric(tmp[,2])), nrow=6, byrow=T)
total_totals <- rowSums(tmp2)
max_tot <- max(total_totals)
most_active <- unique(tmp[,1])[total_totals == max_tot]
print(paste("Most active bird: ",most_active))
```

## Determine which bird is the least active in the Room 3 Dataset

```{r who is the least active?}
room3_daily <- data.frame()
room3_weekly <- data.frame()
room3_monthly <- data.frame()
#TODO select sample as opposed to ID for intra bird comparison (alpha eq)
for(item in trans_reslist_room_3){
  room3_daily <- rbind(room3_daily, item$daily[c("ID","bottom","mid","top","total")])
  room3_weekly <- rbind(room3_weekly, item$weekly[c("ID","bottom","mid","top","total")])
  room3_monthly <- rbind(room3_monthly, item$monthly[c("ID","bottom","mid","top","total")])
}
tmp <- cbind(as.character(room3_daily$ID), as.numeric(room3_daily$total))
#nrow=32 comes from 32 birds used
tmp2 <- matrix(as.vector(as.numeric(tmp[,2])), nrow=6, byrow=T)
total_totals <- rowSums(tmp2)
max_tot <- min(total_totals)
least_active <- unique(tmp[,1])[total_totals == max_tot]
print(paste("Least active bird: ",least_active))
```

# Most Active Bird

```{r most active bird analysis}
set.seed(34716)
library(scales)
library(stringr)
library(ggplot2)
# feb_to_may_daily <- data.frame()
# feb_to_may_weekly <- data.frame()
# feb_to_may_monthly <- data.frame()
# #TODO select sample as opposed to ID for intra bird comparison (alpha eq)
# for(item in trans_reslist_feb){
#   feb_to_may_daily <- rbind(feb_to_may_daily, item$daily[c("ID","bottom","mid","top","total")])
#   feb_to_may_weekly <- rbind(feb_to_may_weekly, item$weekly[c("ID","bottom","mid","top","total")])
#   feb_to_may_monthly <- rbind(feb_to_may_monthly, item$monthly[c("ID","bottom","mid","top","total")])
# }
# tmp <- cbind(as.character(feb_to_may_daily$ID), as.numeric(feb_to_may_daily$total))
# tmp2 <- matrix(as.vector(as.numeric(tmp[,2])), nrow=32, byrow=T)
# total_totals <- rowSums(tmp2)
# max_tot <- max(total_totals)
# most_active <- unique(tmp[,1])[total_totals == max_tot]

room3_daily <- data.frame()
room3_weekly <- data.frame()
room3_monthly <- data.frame()

for(item in trans_reslist_room_3){
  room3_daily <- rbind(room3_daily, item$daily[c("sample","bottom","mid","top","total")])
  room3_weekly <- rbind(room3_weekly, item$weekly[c("sample","bottom","mid","top","total")])
  room3_monthly <- rbind(room3_monthly, item$monthly[c("sample","bottom","mid","top","total")])
}

room3_daily

most_active_intra_day <- room3_daily[grep(most_active,room3_daily$sample),]
most_active_intra_day <- as.data.frame(most_active_intra_day)
most_active_intra_day[,2:5] <- sapply(most_active_intra_day[,2:5], as.numeric)
boxplot(most_active_intra_day[,2:5])
hist(most_active_intra_day[,2])

wss <- 0

for(i in 1:15){
  km.out <- kmeans(most_active_intra_day[,2:4], centers = i, nstart=20)
  wss[i] <- km.out$tot.withinss
}


plot(1:15, wss, type="b")

km.out <- kmeans(most_active_intra_day[,2:4],3, 20)
#km.out <- kmeans(most_active_intra_day[,2:4],3, 20)
summary(km.out)
table(km.out$cluster)

pr.feb <- prcomp(x=most_active_intra_day[,2:4], scale=T, center=T)
pr.feb.prop <- summary(pr.feb)
most_active_intra_day["day"] <- as.numeric(str_split_fixed(most_active_intra_day$sample, "d.",2)[,2])
most_active_intra_day["cluster"] <- km.out$cluster
most_active_intra_day
pr.feb$x <- data.frame(pr.feb$x)

p <- ggplot(pr.feb$x[,1:2], aes(x=PC1, y=PC2, color=as.character(most_active_intra_day$cluster)))+
  geom_point() +
  labs(title=paste("Most Active Bird (",most_active,")"),x=paste("PC1 (",round(pr.feb.prop$importance[[2,1]]*100,2),"%)"),y=paste("PC2 (",round(pr.feb.prop$importance[[2,2]]*100,2),"%)"),color="Cluster") +
  scale_color_discrete(name = "Days", labels= c(as.character(paste(most_active_intra_day$day[most_active_intra_day$cluster == 1],collapse=",")),as.character(paste(most_active_intra_day$day[most_active_intra_day$cluster == 2],collapse=",")),as.character(paste(most_active_intra_day$day[most_active_intra_day$cluster == 3],collapse=",")))) +
  theme_bw() + 
  theme(legend.position = "bottom", legend.direction = "vertical",legend.text = element_text(size=8))
p
ggsave(p, filename = "../figures/most_active_bird_pca_rm3.png",device = "png",width = unit(7,"in"),height=unit(7,"in"))


cluster_table <- data.frame(cbind(as.numeric(km.out$cluster),most_active_intra_day$sample))
cluster_table[order(cluster_table$X1),]

most_active_intra_day[order(most_active_intra_day$total),]



most_active_grp <- data.frame(x = most_active_intra_day$day, y = c(most_active_intra_day$bottom, most_active_intra_day$mid, most_active_intra_day$top), group= c(rep("Bottom",nrow(most_active_intra_day)),rep("Mid",nrow(most_active_intra_day)),rep("Top",nrow(most_active_intra_day))))

 q <- ggplot(most_active_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title=paste("Most Active Bird (",most_active,") Transitions"), color="Legend")+
  scale_x_continuous(name="Days",n.breaks = 14)+
  scale_y_continuous(name="Transitions",n.breaks = 8) +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
q
ggsave(q, filename = "../figures/most_active_transitions_per_day_rm3.png",device="png")
```

# Least Active Bird

```{r least active bird time}
library(stringr)
library(ggplot2)
# feb_to_may_daily <- data.frame()
# feb_to_may_weekly <- data.frame()
# feb_to_may_monthly <- data.frame()
# #TODO select sample as opposed to ID for intra bird comparison (alpha eq)
# for(item in trans_reslist_feb){
#   feb_to_may_daily <- rbind(feb_to_may_daily, item$daily[c("ID","bottom","mid","top","total")])
#   feb_to_may_weekly <- rbind(feb_to_may_weekly, item$weekly[c("ID","bottom","mid","top","total")])
#   feb_to_may_monthly <- rbind(feb_to_may_monthly, item$monthly[c("ID","bottom","mid","top","total")])
# }
# 
# 
# tmp <- cbind(as.character(feb_to_may_daily$ID), as.numeric(feb_to_may_daily$total))
# tmp2 <- matrix(as.vector(as.numeric(tmp[,2])), nrow=32, byrow=T)
# total_totals <- rowSums(tmp2)
# min_tot <- min(total_totals)
# least_active <- unique(tmp[,1])[total_totals == min_tot]

room3_daily <- data.frame()
room3_weekly <- data.frame()
room3_monthly <- data.frame()
#TODO select sample as opposed to ID for intra bird comparison (alpha eq)
for(item in trans_reslist_room_3){
  room3_daily <- rbind(room3_daily, item$daily[c("sample","bottom","mid","top","total")])
  room3_weekly <- rbind(room3_weekly, item$weekly[c("sample","bottom","mid","top","total")])
  room3_monthly <- rbind(room3_monthly, item$monthly[c("sample","bottom","mid","top","total")])
}

room3_daily

least_active_intra_day <- room3_daily[grep(least_active,room3_daily$sample),]
least_active_intra_day <- as.data.frame(least_active_intra_day)
least_active_intra_day[,2:5] <- sapply(least_active_intra_day[,2:5], as.numeric)
boxplot(least_active_intra_day[,2:5])
hist(least_active_intra_day[,4])

wss <- 0

for(i in 1:15){
  km.out <- kmeans(least_active_intra_day[,2:4], centers = i, nstart=20)
  wss[i] <- km.out$tot.withinss
}


plot(1:15, wss, type="b")

km.out <- kmeans(least_active_intra_day[,2:4],3, 20)
summary(km.out)
table(km.out$cluster)

pr.feb <- prcomp(x=least_active_intra_day[,2:4], scale=T, center=F)
pr.feb.prop <- summary(pr.feb)

pr.feb.prop$importance
pr.feb$x <- data.frame(pr.feb$x)


least_active_intra_day["day"] <- as.numeric(str_split_fixed(least_active_intra_day$sample, "d.",2)[,2])
least_active_intra_day["cluster"] <- km.out$cluster
least_active_intra_day

p <- ggplot(pr.feb$x[,1:2], aes(x=PC1, y=PC2, color=as.character(least_active_intra_day$cluster)))+
  geom_point() + 
  labs(title=paste("Least Active Bird (",least_active,")"),x=paste("PC1 (",round(pr.feb.prop$importance[[2,1]]*100,2),"%)"),y=paste("PC2 (",round(pr.feb.prop$importance[[2,2]]*100,2),"%)"),color="Cluster") +
  scale_color_discrete(name = "Days", labels= c(as.character(paste(least_active_intra_day$day[least_active_intra_day$cluster == 1],collapse=",")),as.character(paste(least_active_intra_day$day[least_active_intra_day$cluster == 2],collapse=",")),as.character(paste(least_active_intra_day$day[least_active_intra_day$cluster == 3],collapse=",")))) +
  theme_bw() + 
  theme(legend.position = "bottom", legend.direction = "vertical",legend.text = element_text(size=8))
p
ggsave(p, filename = "../figures/least_active_bird_pca_rm3.png",device = "png",width = unit(7.5,"in"),height=unit(7.5,"in"))

cluster_table <- data.frame(cbind(as.numeric(km.out$cluster),least_active_intra_day$sample))
cluster_table[order(cluster_table$X1),]

least_active_intra_day[order(least_active_intra_day$total),]
least_active_intra_day




least_active_grp <- data.frame(x = least_active_intra_day$day, y = c(least_active_intra_day$bottom, least_active_intra_day$mid, least_active_intra_day$top), group= c(rep("Bottom",nrow(least_active_intra_day)),rep("Mid",nrow(least_active_intra_day)),rep("Top",nrow(least_active_intra_day))))



q <-ggplot(least_active_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title=paste("Least Active Bird (",least_active,") Transitions"), color="Legend")+
  scale_x_continuous(name="Days",n.breaks = 14)+
  scale_y_continuous(name="Transitions",n.breaks = 8) +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
q
ggsave(q, filename="../figures/least_active_transitions_per_day_rm3.png",device="png")
```

# Organize Birds By activity 

```{r evaluate transition clolumn sum on the daily data}
library(qpcR)
room3_daily_colsum <- data.frame()

#each item is one transition list for an intra bird sample
#we want to sum up the columns for each day and compare bird to bird


for(i in 1:length(trans_reslist_room_3)){
  res <- colSums(sapply(trans_reslist_room_3[[i]]$daily[,2:6], as.numeric))
  res[1] <- res[1]/length(trans_reslist_room_3[[i]]$daily$ID)
  res[6] <- round(res[2]/length(trans_reslist_room_3[[i]]$daily$bottom),0)
  res[7] <- round(res[3]/length(trans_reslist_room_3[[i]]$daily$mid),0)
  res[8] <- round(res[4]/length(trans_reslist_room_3[[i]]$daily$top),0)
  res[9] <- round(res[5]/length(trans_reslist_room_3[[i]]$daily$total),0)
  room3_daily_colsum <- rbind(room3_daily_colsum,res)
}

colnames(room3_daily_colsum) <- c("ID","bottom","mid","top","total","bottom_A","mid_A","top_a","total_A")
room3_daily_colsum
hold <- room3_daily_colsum[order(room3_daily_colsum$total_A,decreasing = T),]

summary(hold[,6:9])
boxplot(hold[,6:9])

low_act <- c()
med_act <- c()
high_act <- c()

for(i in 1:length(hold[,1])){
  cur_tol <- hold[i,9]
  if(cur_tol<=50){
    low_act <- c(low_act,hold[i,1])
  }
  if((cur_tol>50) & (cur_tol<=110)){
    med_act <- c(med_act,hold[i,1])
  }
  if(cur_tol>110){
    high_act <- c(high_act, hold[i,1])
  }
}


cluster_table <- data.frame(qpcR:::cbind.na(low_act,med_act,high_act))
cluster_table
write.csv(cluster_table, "../output/organized_birds_by_activity_daily_rm3.csv",row.names = F)
```

## Jumpy table for the separated samples

```{r make a jumpy table for the low activity birds}
# Aggregate the low activity birds
low_act_bird_ids <- na.trim(cluster_table$low_act)

#make a table of the samples
low_act_bird_frame <- data.frame()

for(id in low_act_bird_ids){
  curr_bird <- room3_daily[grep(id,room3_daily$sample),]
  low_act_bird_frame <- rbind(low_act_bird_frame,curr_bird)
}

low_act_bird_frame[,2:5] <- sapply(low_act_bird_frame[,2:5], as.numeric)
low_act_bird_frame["day"] <- as.numeric(str_split_fixed(low_act_bird_frame$sample, "d.",2)[,2])
low_act_agg_tab <- aggregate(low_act_bird_frame[,2:5],by = list(low_act_bird_frame$day), FUN = mean)
colnames(low_act_agg_tab) <- c("day","bottom","mid","top","total")

# plot the low activity birds
low_act_grp <- data.frame(x = low_act_agg_tab$day, y = c(low_act_agg_tab$bottom, low_act_agg_tab$mid, low_act_agg_tab$top), group= c(rep("Bottom",nrow(low_act_agg_tab)),rep("Mid",nrow(low_act_agg_tab)),rep("Top",nrow(low_act_agg_tab))))



q <-ggplot(low_act_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title="Low Activity Birds Transitions in Room 3", color="Legend")+
  scale_x_continuous(name="Days",n.breaks = 14)+
  scale_y_continuous(name="Transitions") +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
q
print("which birds are in this plot?")
print(low_act_bird_ids)
ggsave(q, filename="../figures/low_active_transitions_rm3.png",device="png")
```

```{r make a jumpy table for the medium activity birds}
# Aggregate the medium activity birds
med_act_bird_ids <- na.trim(cluster_table$med_act)

#make a table of the samples
med_act_bird_frame <- data.frame()

for(id in med_act_bird_ids){
  curr_bird <- room3_daily[grep(id,room3_daily$sample),]
  med_act_bird_frame <- rbind(med_act_bird_frame,curr_bird)
}

med_act_bird_frame[,2:5] <- sapply(med_act_bird_frame[,2:5], as.numeric)
med_act_bird_frame["day"] <- as.numeric(str_split_fixed(med_act_bird_frame$sample, "d.",2)[,2])
med_act_agg_tab <- aggregate(med_act_bird_frame[,2:5],by = list(med_act_bird_frame$day), FUN = mean)
colnames(med_act_agg_tab) <- c("day","bottom","mid","top","total")

# plot the low activity birds
med_act_grp <- data.frame(x = med_act_agg_tab$day, y = c(med_act_agg_tab$bottom, med_act_agg_tab$mid, med_act_agg_tab$top), group= c(rep("Bottom",nrow(med_act_agg_tab)),rep("Mid",nrow(med_act_agg_tab)),rep("Top",nrow(med_act_agg_tab))))



q <-ggplot(med_act_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title="Medium Activity Birds Transitions in Room 2", color="Legend")+
  scale_x_continuous(name="Days",n.breaks = 14)+
  scale_y_continuous(name="Transitions") +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
q
print("which birds are in this plot?")
print(med_act_bird_ids)
ggsave(q, filename="../figures/med_active_transitions_rm3.png",device="png")
```

```{r make a jumpy table for the high activity birds}
# Aggregate the high activity birds
high_act_bird_ids <- na.trim(cluster_table$high_act)

#make a table of the samples
high_act_bird_frame <- data.frame()

for(id in high_act_bird_ids){
  curr_bird <- room3_daily[grep(id,room3_daily$sample),]
  high_act_bird_frame <- rbind(high_act_bird_frame,curr_bird)
}

high_act_bird_frame[,2:5] <- sapply(high_act_bird_frame[,2:5], as.numeric)
high_act_bird_frame["day"] <- as.numeric(str_split_fixed(high_act_bird_frame$sample, "d.",2)[,2])
high_act_agg_tab <- aggregate(high_act_bird_frame[,2:5],by = list(high_act_bird_frame$day), FUN = mean)
colnames(high_act_agg_tab) <- c("day","bottom","mid","top","total")

# plot the low activity birds
high_act_grp <- data.frame(x = high_act_agg_tab$day, y = c(high_act_agg_tab$bottom, high_act_agg_tab$mid, high_act_agg_tab$top), group= c(rep("Bottom",nrow(high_act_agg_tab)),rep("Mid",nrow(high_act_agg_tab)),rep("Top",nrow(high_act_agg_tab))))



q <-ggplot(high_act_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title="High Activity Birds Transitions in Room 3", color="Legend")+
  scale_x_continuous(name="Days",n.breaks = 14)+
  scale_y_continuous(name="Transitions") +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
q
print("which birds are in this plot?")
print(high_act_bird_ids)
ggsave(q, filename="../figures/high_active_transitions_rm3.png",device="png")
```

# How long does the bird stay in a zone?

```{r how long does the bird stay in the zone}

trans_reslist_room3_min <- list()

for(i in 1:length(reslist2_9013)){
  minute_period <- to_minute(reslist2_9013[[i]]$xts_obj$subzone)
  minute_trans_table <- calc_trans_period(reslist2_9013[[i]]$ID, minute_period, reslist2_9013[[i]]$xts_obj, "Hr")  
  result <- list("ID"=reslist2_9013[[i]]$ID,"hour"=minute_trans_table, "raw"=reslist2_9013[[i]]$xts_obj)
  trans_reslist_room3_min[[i]] <- result
}

room3_hour <- data.frame()

for(item in trans_reslist_room3_min){
  room3_hour <- rbind(room3_hour, item$hour[c("sample","bottom","mid","top","total")])
}

# check to make sure only one bird is selected
print("check that the id and RFID tag are unique for each item entry")
#this is following up with the issues of the intra-bird comparisons
for(item in trans_reslist_room3_min){
  print(paste(unique(item$raw$Tagnumber)," ",unique(item$raw$tagname)))
  #print(item)
}

room3_hourly <- data.frame()

#each item is one transition list for an intra bird sample
#we want to sum up the columns for each day and compare bird to bird

room3_hourly_colsum_min <- data.frame()

for(i in 1:length(trans_reslist_room3_min)){
  res <- colSums(sapply(trans_reslist_room3_min[[i]]$hour[,2:6], as.numeric))
  res[1] <- res[1]/length(trans_reslist_room3_min[[i]]$hour$ID)
  res[6] <- round(res[2]/length(trans_reslist_room3_min[[i]]$hour$bottom),0)
  res[7] <- round(res[3]/length(trans_reslist_room3_min[[i]]$hour$mid),0)
  res[8] <- round(res[4]/length(trans_reslist_room3_min[[i]]$hour$top),0)
  res[9] <- round(res[5]/length(trans_reslist_room3_min[[i]]$hour$total),0)
  room3_hourly_colsum_min <- rbind(room3_hourly_colsum_min,res)
}

colnames(room3_hourly_colsum_min) <- c("ID","bottom","mid","top","total","bottom_A","mid_A","top_a","total_A")
room3_hourly_colsum_min
hold_hour <- room3_hourly_colsum_min[order(room3_hourly_colsum_min$total,decreasing = T),]

summary(hold_hour[,6:9])
boxplot(hold_hour[,6:9])

hold_hour

```


```{r calculate duration traslist}

trans_reslist_rm_3 <- c()
minute_reslist_room3_hr <- c()

for(i in 1:length(reslist2_9013)){
  minute_period <- to_minute(reslist2_9013[[i]]$xts_obj$subzone)
  minute_reslist_room3_hr[[i]] <- minute_period
}

trans_reslist_room3_dur <- c()
for(i in 1:length(minute_reslist_room3_hr)){
  minute_trans_tab <- calc_trans_period_duration(reslist2_9013[[i]]$ID, minute_reslist_room3_hr[[i]],reslist2_9013[[i]]$xts_obj, "hr")
  #minute_trans_table <- calc_trans_period_duration(reslist2_6928[[i]]$ID, minute_period, reslist2_6928[[i]]$xts_obj, "Hr")
  result <- list("ID"=reslist2_9013[[i]]$ID,"hour"=minute_trans_tab,"raw"=reslist2_9013[[i]]$xts_obj)
  trans_reslist_room3_dur[[i]] <- result
}

#trans_reslist_feb_dur
```



```{r make a jumpy table that looks at duration}

#which bird is this?

trans_reslist_room3_dur[[1]]$ID

# Aggregate the high activity birds
example_duration <- na.trim(trans_reslist_room3_dur[[1]]$hour)

#make a table of the samples
duration_frame <- data.frame(example_duration)

duration_frame[,3:9] <- sapply(duration_frame[,3:9], as.numeric)
duration_frame["hour"] <- as.numeric(str_split_fixed(duration_frame$sample, "hr.",2)[,2])
duration_frame["day"] <- as.integer((duration_frame$hour/24)+2)
duration_agg_tab_hr <- aggregate(duration_frame[,7:9],by = list(duration_frame$hour), FUN = sum)
colnames(duration_agg_tab_hr) <- c("hour","bottom","mid","top")

duration_agg_tab_day <- aggregate(duration_frame[,7:9],by = list(duration_frame$day), FUN = sum)
colnames(duration_agg_tab_day) <- c("day","bottom","mid","top")

# plot the low activity birds
duration_hr_grp <- data.frame(x = duration_agg_tab_hr$hour, y = c(duration_agg_tab_hr$bottom, duration_agg_tab_hr$mid, duration_agg_tab_hr$top), group= c(rep("Bottom",nrow(duration_agg_tab_hr)),rep("Mid",nrow(duration_agg_tab_hr)),rep("Top",nrow(duration_agg_tab_hr))))

duration_day_grp <- data.frame(x = duration_agg_tab_day$day, y = c(duration_agg_tab_day$bottom, duration_agg_tab_day$mid, duration_agg_tab_day$top), group= c(rep("Bottom",nrow(duration_agg_tab_day)),rep("Mid",nrow(duration_agg_tab_day)),rep("Top",nrow(duration_agg_tab_day))))


q <-ggplot(duration_hr_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title="Example Birds Transition Duration in Room 3", color="Legend")+
  scale_x_continuous(name="Hours",n.breaks = 100)+
  scale_y_continuous(name="Seconds Spent in Zone") +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
q

z <-ggplot(duration_day_grp, aes(x,y,col=group))+
  geom_line() +
  facet_grid(factor(group,levels=c("Top","Mid","Bottom"))~ .) +
  labs(title="Example Birds Transition Duration in Room 3", color="Legend")+
  scale_x_continuous(name="Days",n.breaks = 14)+
  scale_y_continuous(name="Seconds Spent in Zone") +
  scale_color_manual(values=c("Top"="royalblue","Mid"="tomato","Bottom" = "seagreen")) +
  theme_bw()
z


ggsave(q, filename="../figures/duration_transitions_rm3.png",device="png")
ggsave(z, filename="../figures/duration_transitions_day_rm3.png",device="png")
```
